---
title: "Naming Things"
description: |
  A brief introduction to best practices for file and directory naming that will help keep you sane.
preview: files/feature.png
#author:
#  - name: "Jarrod"
#    orcid_id: 0000-0001-9863-1318
#    url: https://metacrobe.github.io/
#    affiliation: Smithsonian Tropical Research Institute
#    affiliation_url: https://stri.si.edu/
date: 2022-04-21
output:
  distill::distill_article:
    self_contained: false
    toc: false
    toc_depth: 2
    #css: ../../assets/styles.css
bibliography: ../../assets/cite.bib  
repository_url: "https://github.com/stri-mcgill-neo/2022/"
#collections:
#  posts:
#    disqus: reproducible-finance-with-r
#    share: [twitter, linkedin]
#    citations: false
draft: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
is_html = knitr::is_html_output()
pacman::p_load(tidyverse, fontawesome, slickR,
               install = FALSE, update = FALSE)
source(file.path("../../assets", "functions.R"))
```

```{r echo=FALSE, eval=TRUE}
# This is needed to add a COPY CODE button to code chunks
clipboard_button()
# This is needed to use PANELS
xaringanExtra::use_panelset()
```

```{r, xaringanExtra-clipboard, echo=FALSE, eval=TRUE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa fa-clone fa-2x\" style=\"color: #004276\"></i>",
    success_text = "<i class=\"fa fa-check fa-2x\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times fa-2x\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
```

This is a  brief note on *best practices* for naming files and directories. There is no universal convention that I am aware of but you can save yourself a lot of trouble by following a few simple rules when naming files and directories.

<aside>
Estimated reading time: `r round(reading_time(), digits = 0)` minute(s).
</aside>


> Your task is to ensure your files are **Machine readable** and **Human readable**.

**Never** use special characters in your file name (e.g., `* # : \ / < > | " ? [ ] ; = + & £ $`,  and so on). Letters and numbers **only**. Many of these symbols are prohibited and a lot of software will reject files that contain these symbols. 

Periods `.` should only be used for a file extension, like `document.html`. If you put a period `.` at the beginning of a file or directory name, like `.bash_profile`, then it will be hidden by default. Do not do this at the beginning of a name unless you know what you are doing :) 

**Do not** use spaces. Separate words using a dash (`-`) or an underscore (`_`), like `how-to-use-this-site.Rmd`. You can also use camel case if you prefer, like `howToUseThisSite.Rmd`.

The file name should be short but meaningful. This is important when dealing with a lot of files, like for a website. So add relevant details but keep it less than 25 characters---long file names mean long URLs, which can increase the likelihood of errors.. 

If you include a date in the name, use date format [ISO_8601](https://en.wikipedia.org/wiki/ISO_8601) `YYYY-MM-DD` or `YYYYMMDD`. Distill blog sites use `YYYY-MM-DD` by default. 

You can begin a file/directory name with a number, just be sure to use sequential numbering like `01_File`, `02_Stuff`, etc. instead of `1_File` and `2_Stuff`.

Whatever approach you choose, I think the main key is to be consistent. If you use dashes `-` to separate words in a name, then try to stick with that for everything.  

## Credits {.appendix}

The feature photo for this post---[A small amateur insect collection by](https://commons.wikimedia.org/wiki/File:Small-insect-collection.jpg) [Debivort](https://en.wikipedia.org/wiki/User:Debivort) at the English-language Wikipedia---was retrieved from Wikimedia Commons and Licensed under [CC BY-SA 3.0](http://creativecommons.org/licenses/by-sa/3.0/). Original image found [here](https://upload.wikimedia.org/wikipedia/commons/1/1f/Small-insect-collection.jpg).

##  Source Code {.appendix}

The source code for this page can be accessed on GitHub `r fa(name = "github")` by [clicking this link](`r source_code()`). 
