---
title: "Showcase"
description: |
  A showcase of R Markdown Projects. If you have a project you would like to share, let me know and I will post it here!
output:
  distill::distill_article:
    toc: true
    #keep_md: true
base_url: https://stri-mcgill-neo.github.io/2022/
preview: files/panama_logo.png
params:
  update_all: FALSE
  check_url: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
# Load libraries and functions
source(file.path("files", "functions.R"))
```

This page^[The inspiration for this page comes from [the Distillery showcase page](https://distillery.rbind.io/showcase.html). The code was harvested and modified from [the Distillery GitHub repop](https://github.com/jhelvy/distillery)] is currently just a collection of some web products I have created with R Markdown. It is not that I think these are particularly amazing, it is just that since I coded everything, I know how all of this works. So if you are looking for ideas and inspiration have a look around. 

If you would like to share something you have created, let me know and I will post it on this page :)

```{r build-content, include=FALSE}
sites <- clean_sites(read_csv("files/showcase/sites.csv"), params$check_url)
#update_screenshots(sites, params$update_all)
rmd_chunks <- make_showcase_chunks(sites, image_width = 600)
button_filters <- buttons_filter(sites)
```

Sites and source code are linked on the right of each image. Use the buttons to filter the showcase by category:

```{r insert-showcase, child = c(button_filters, rmd_chunks)}
```


<script src = "assets/js/filter-buttons.js"></script>