---
title: "Showcase"
description: |
  A showcase of R Markdown Projects. If you have a project you would like to share, let me know and I will post it here!
output:
  distill::distill_article:
#    theme:
#      version: 4
    toc: false
    css: [assets/css/b4cards.css, assets/css/cards.css]
base_url: https://stri-mcgill-neo.github.io/2022/
preview: files/panama_logo.png
params:
  update_all: FALSE
  check_url: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
# Load libraries and functions
source(file.path("assets", "functions.R"))
library(distilltools)
library(tidyverse)
library(fontawesome)
```

This page^[The inspiration for this page comes from [the Distillery showcase page](https://distillery.rbind.io/showcase.html) and [John Paul Helveston personal website](https://www.jhelvy.com/projects.html). I used the [bs4cards](https://github.com/djnavarro/bs4cards) R package for the gallery. The code was harvested and modified from [jhelvy\/jhelvy_distill GitHub repo](https://github.com/jhelvy/jhelvy_distill/blob/master/projects.Rmd)] is currently just a collection of some web products I have created with R Markdown. It is not that I think these are particularly amazing, it is just that since I coded everything, I know how all of this works. So if you are looking for ideas and inspiration have a look around. 

If you would like to share something you have created, please let me know---I would love to add it to this page. Use the buttons to filter the showcase by category:

</br>

```{r build-content, include=FALSE}
# Read in projects csv file 
projects <- readr::read_csv(here::here("files/showcase", "bs4cards.csv"))
```


```{r projects, echo=FALSE, layout="l-body-outset"}
projects %>% 
  bs4cards::cards(
    title = title,
    text = description,
    link = url,
    image = src,
    label_colour = "#ffffff",
    tags = paste("Show all;", tags),
    footer = tags,
    border_radius = 5
  )
```


<script src = "assets/js/escape-html.js"></script>

##  Source Code {.appendix}

The source code for this page can be accessed on GitHub `r fa(name = "github")` by [clicking this link](`r source_code()`). 
