---
title: "Tutorials"
#author:
#- name: Jarrod
#  url: https://metacrobe.github.io
date: \`r format(Sys.Date(), "%B %d, %Y")`\
description: |
  A quick reference for key elements used on this site; designed to help make the information more accessible and help you learn more effectively. 
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
library(fontawesome)
knitr::opts_chunk$set(echo = FALSE)
```

```{r, xaringanExtra-clipboard, echo=FALSE, eval=TRUE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa fa-clone fa-2x\" style=\"color: #4682B4\"></i>",
    success_text = "<i class=\"fa fa-check fa-2x\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times fa-2x\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
```

```{r echo=FALSE, eval=TRUE}
ldir <-  normalizePath(".")
finf <- file.info(paste(ldir, "about.Rmd", sep = "/"), extra_cols = FALSE)

bytes <- finf$size
words <- bytes/10
minutes <- words/200
```

## Setting up git/GitHub

1. Sign up for a GitHub account
2. make sure git is installed
3. ssh-keygen -t rsa -b 4096 -C "<your email address>"
4. ls -al .ssh/
5. git config --global user.name "<your name>"
6. git config --global user.email "<your email address>"
7. cd .ssh/
8. nano config
Host *
 AddKeysToAgent yes
 UseKeychain yes
 IdentityFile ~/.ssh/id_rsa
 
9. cd ../
10. ssh-add -K ~/.ssh/id_rsa
Identity added: /Users/rad/.ssh/id_rsa (jarrod.jude.scott@gmail.com)
11. pbcopy < ~/.ssh/id_rsa.pub
12. GitHub < Settings < SSH and GPG Keys
13. New SSH Key
14. give it a unique name
15. paste key in KEY




## Command Line Interface

A command line interface (CLI) is a text-based user interface (UI) used to view and manage computer files. This is in contrast to the graphical user interface (GUI) that most of us are accustomed to. All of the pointing, clicking, copying, and pasting you do with your computer's GUI can be done in the command line. Think of the CLI as the layer just below the GUI.

Knowing how to use the CLI is important for a lot of [reasons](https://www.dataquest.io/blog/why-learn-the-command-line/) (#1 does not apply to this course). For one, using git and GitHub from the command line is far easier and more useful than the git GUI programs.

Ok, here is where I get into trouble. Since I do not have a Windows machine and have so little experience with Windows CLI (DOS), I cannot vet any tutorial the same way I would for other operating systems. I would really appreciate Windows users recommending some good tutorials. I think the syntax and commands are [the same or very similar](https://www.w3schools.com/whatis/whatis_cli.asp).

<i class="fa fa-exclamation-triangle fa-1x" style="color:#006DDB;"></i> A word of caution. The command `rm` stands for **remove**. It is similar to moving a file to the trash except `rm` *deletes* the file. It is extremely difficult (impossible?) to undo `rm` so it is good practice to always include the `-i` flag when running this command. Adding this flag forces, you to confirm that you want to delete the file. If I want to remove a file called `file.txt` I would run the command like this: `rm -i file.txt`. Once I hit enter, I receive a message that asks `remove file.txt?`. I must answer `y` or `n` or the command will not run.

Many important files and directories on your computer are **write protected**, meaning that you must enter a password to delete them. Most of the time you will be ok. But running `rm` incorrectly, without the `-i` flag, in the wrong place on your computer, can be **catastrophic**. Please be careful with your new powers.

### CLI Tutorials

* [Conquering the Command Line](http://conqueringthecommandline.com/book/basics). Excellent and extensive tutorials from Mark Bates.
* [Briana's Bash Tutorial: How to Use the Command Line](https://www.youtube.com/watch?v=BFMyUgF6I8Y). Twenty minute video from freeCodeCamp.org.
* [Beginner's Guide to the Bash Terminal](https://www.youtube.com/watch?v=oxuRxtrO2Ag). Longer video from [Joe Collins](https://www.youtube.com/channel/UCTfabOKD7Yty6sDF4POBVqA). Amazing.

## HTML & CSS

Websites are built on HTML, CSS, and JavaScript. For any site, HTML defines the content and structure; CSS provides the style and layout; JavaScript programs the behavior.

If you want to develop your HTML/CSS skills, you can start by working through these tutorials. At this point in the course, learning HTML/CSS is ~~*optional*~~ self-paced, but if there is enough demand for more, I can incorporate these languages more formerly. If you have no interest in learning these languages that is totally fine. You do not need them to use R Markdown. I do believe however that a little knowledge of HTML and CSS goes a long way. As I mentioned, if you go far enough with R Markdown, you will encounter its limitations. I encourage you check out some of these tutorials.

### Interneting is Hard, But is doesn't have to be

I have explored a lot of tutorials for HTML/CSS and this is my favorite. There are 14 lessons that cover  many important skills. The tutorials are simple and straightforward. It is also a good opportunity to start working in your fancy new source code editor and testing out the dev tool functions of your browser.

[A friendly web development tutorial for complete beginners](https://www.internetingishard.com/html-and-css/).

***

## Additional Tutorials

In this section, I will add other tutorials you can explore on your own time for honing your skills. If you find a tutorial you like, please let me know and I will add it to the list.

## General

[Learn X in Y minutes](https://learnxinyminutes.com/). Where `X` is a programming language. Beginner tutorials available English, Spanish, and many other languages. Use the *Highlight your language:* button in the upper right corner to filter. I highlight some in the sections below.

[80+ Ways to Learn to Code for Free in 2020](https://skillcrush.com/blog/64-online-resources-to-learn-to-code-for-free/). 

### HTML/CSS

From Learn X in Y minutes, [CSS in English (EN)](https://learnxinyminutes.com/docs/css/) (no ES CSS available), [HTML in English (EN)](https://learnxinyminutes.com/docs/html/), and [HTML in Espanol (ES)](https://learnxinyminutes.com/docs/es-es/html-es/).

Both w3schools and MDN web docs have extensive information about CSS and HTML. These are amazing resources.
* [w3schools HTML](https://www.w3schools.com/html/default.asp). HTML tutorials, examples, and explanations.
* [w3schools CSS](https://www.w3schools.com/css/default.asp). CSS tutorials, examples, and explanations.
* [MDN web docs for HTML](https://developer.mozilla.org/en-US/docs/Web/HTML). HTML tutorials, examples, and explanations.
* [MDN web docs for CSS](https://developer.mozilla.org/en-US/docs/Web/CSS).  CSS tutorials, examples, and explanations.
* MDN web docs also has many tutorials in [Spanish](https://developer.mozilla.org/es/docs/Web/Tutoriales).

### Markdown

From Learn X in Y minutes, [Markdown in English (EN)](https://learnxinyminutes.com/docs/markdown/) and [Markdown en Espanol (ES)](https://learnxinyminutes.com/docs/es-es/markdown-es/). Here is a

### R

From Learn X in Y minutes, [R in English (EN)](https://learnxinyminutes.com/docs/r/) and [R en Espanol (ES)](https://learnxinyminutes.com/docs/es-es/r-es/).

### MathJax

* [Basic tutorial & quick reference](https://math.meta.stackexchange.com/questions/5020/mathjax-basic-tutorial-and-quick-reference) from StackExchange. A bit old now but still good.


{{< edit "https://github.com/stri-con/data-curation/blob/master/content/page/tutorials.md" >}}


